---
title: "4.21_G_computation_simulation_glinternet"
Author: "Yulin Gong"
format:
  html:
    theme: cosmo
    embed-resources: true
    number-sections: false
    toc: true
    toc-depth: 4
    toc-location: left
    code-fold: false
---

# Load packages and functions
```{r}



if (!require("pacman")){
  install.packages("pacman", repos = 'http://cran.us.r-project.org')
}

pacman::p_load("glmnet",
               "tidyverse",
               "rio",
               "magrittr",
               "broom",
               "labelled",
               "Hmisc",
               "flextable",
               "gtsummary",
               "emmeans",
               "here",
               "cleaR",
               "fastDummies",
               "formattable",
               "furrr",
               "future",
               "glue")
```



```{r}
cleaR::clear()
{
  
    rmulti = function(prob, n) {
    as.numeric(rMultinom(probs=matrix(prob, 1, length(prob)), m=n))
  }
  
  expit <- function(linearcombination){
    n = (1/(1 + exp(-(linearcombination))))
    return (n)
  }
  
  logit <- function(pr){
    n = log(pr/(1-pr))
    return(n)
  }
  
round_df <- function(df, digits) {
  nums <- vapply(df, is.numeric, FUN.VALUE = logical(1))
  
  df[,nums] <- round(df[,nums], digits = digits)
  (df)
}

diff_fct <- function(x){
  dif <- x - lag(x)
  return(dif)
}


p_load("rlang")
recreate_var <- function(datau, var_name) {
  
  vars <- strsplit(var_name, "__")[[1]]
  
  expr <- rlang::parse_expr(paste(vars, collapse = " * "))
  
  datau %>% mutate("{var_name}" := !!expr)
}

geom_hist_percent <- function(.data, .var) {
  .data %>%
    ggplot(aes(x = !!sym(.var))) +
    geom_histogram(aes(y = after_stat(count/sum(count)))) +
    ylab("Percent") +
    scale_y_continuous(labels = scales::percent) +
    theme_bw()
}
}
```

# GENERALIZED HTE APPROACH (SIMULATED DATA, 10,000 A1)
## Data generating mechanism for simulation

```{r}

n=10000
set.seed(123)

data_grf <- tibble(
  age   = rnorm(n, mean=68, sd=10), 
  black = rbinom(n, size=1, prob=0.3), 
  sbp   = rnorm(n, mean=140, sd=15), 
  dbp   = rnorm(n, mean=78, sd=12), 
  scr   = rnorm(n, mean=1.07, sd=0.34), 
  egfr  = rnorm(n, mean=72, sd=20), 
  egfr_ge72 = as.numeric(egfr>=72),
  statin=rbinom(n, size=1, prob=0.43), 
  aspirin=rbinom(n, size=1, prob=0.51), 
  fram  = rnorm(n, mean=25, sd=12), 
  smok  = factor(rmulti(prob=c(0.44, 0.42, 0.14), n))) %>% 
  dummy_cols("smok", remove_selected_columns = F,
             remove_first_dummy = T) %>% 
  mutate(
    tx = rbinom(n, size=1, prob=expit(logit(0.01) + 
                                        log(0.90)*statin+
                                        log(1.01)*sbp+
                                        log(1.01)*dbp+
                                        log(1.01)*age+
                                        log(1.03)*fram+
                                        log(1.1)*black+
                                        log(1.1)*scr+
                                        log(1.1)*smok_3 +
                                        log(1.05)*smok_2)),
    
    y = rbinom(n, size=1, prob=expit(logit(0.01) + 
                                       log(0.038)*tx + 
                                       log(0.592)*egfr_ge72 +
                                       log(0.453)*aspirin +
                                       log(12.00)*tx*egfr_ge72 + 
                                       log(34.70)*tx*aspirin +
                                       log(0.543)*egfr_ge72*aspirin + 
                                       log(0.372)*tx*egfr_ge72*aspirin +
                                       
                                       log(0.90)*statin+
                                       log(1.01)*sbp+
                                       log(1.01)*dbp+
                                       log(1.01)*age+
                                       log(1.03)*fram+
                                       log(1.1)*black+
                                       log(1.2)*scr+
                                       log(1.05)*smok_3 +
                                       log(1.05)*smok_2))) %>% 
  set_variable_labels(egfr    = "Estimated Glomerular Filtration Rate",
                      aspirin = "Aspirin use",
                      statin  = "Statin use",
                      sbp     = "Systolic blood pressure",
                      dbp     = "Diastolic blood pressure",
                      age     = "Age in years",
                      fram    = "Framingham score risk",
                      black   = "Black race",
                      scr     = "Serum Creatinine",
                      smok_3  = "Current smoker", #smok_1: "Smoking (never)"
                      smok_2  = "Fomer smoker") 

```

```{r}
library(gtsummary)
library(dplyr)
library(labelled)
library(forcats)

table1 <- data_grf %>% 
  mutate(tx = case_when(
    tx == 1 ~ "Treatment",
    TRUE    ~ "Placebo"
  )) %>%
  mutate(`Smoking status` = as.numeric(as_factor(smok))) %>%  
  dplyr::select(-contains("smok_"), -egfr_ge72, -smok) %>%
  relocate(y, .after = "Smoking status") %>%
  rename(`Y (cardiovascular events)` = y) %>%
  set_value_labels(
    `Smoking status` = c(
      "Never smoker"   = 1,
      "Former smoker"  = 2,
      "Current smoker" = 3
    )
  ) %>%
  modify_if(is.labelled, to_factor) %>%
  tbl_summary(
    missing = "no",
    by = "tx",
    digits = list(everything() ~ 0),
    statistic = list(
      all_continuous()  ~ "{mean} ({sd})",
      all_categorical() ~ "{n} ({p}%)"
    )
  ) %>%
  add_overall() %>%
  add_stat_label(location = "row") %>%
  add_difference(
    test = list(
      all_continuous()  ~ "smd",
      all_categorical() ~ "smd"
    )
  ) %>%
  modify_header(
    label = "**Variable**",
    estimate = "**SMD**"
  ) %>%
  modify_spanning_header(starts_with("stat_") ~ "Treatment Assignment") %>%
  bold_labels()
table1
```

# Step 1: select variables and data preprocessing
```{r}
data_grf <- data_grf %>%
  mutate(
    smok = factor(smok,
                  levels = c(1, 2, 3),
                  labels = c("Never", "Former", "Current"))
  )

dsn1 <- data_grf %>% 
  dplyr::select(-c(egfr_ge72, smok)) # removing these variables to run the analysis as they are no longer needed

exposure ="tx"
outcome  ="y"

covariates = setdiff(colnames(dsn1), c(outcome)) #include exposure here
covariates
```



```{r}

pacman::p_load("WeightIt", "cobalt", "dplyr", "glue", "tidyr", "glinternet", "rsample")

# Note: train/test split is performed later after creating `dsn2`.

# Step 2: Fit a flexible saturated model
# Step 3: Apply a selection method to reduce overfitting



dsn2 <- dsn1 %>%
  mutate(
    smok = dplyr::case_when(
      smok_3 == 1 ~ 3L,
      smok_2 == 1 ~ 2L,
      TRUE        ~ 1L
    ),
    smok = factor(smok, levels = 1:3, labels = c("Never", "Former", "Current"))
  ) %>%
  select(-smok_2, -smok_3)  


data_split <-initial_split(dsn2, 
                           strata = "y", 
                           prop = 0.70)
train <- training(data_split)
test <- testing(data_split)


form_denom <- tx ~ age + black + sbp + dbp + scr + egfr + statin + aspirin + 
    fram + smok
# Estimate propensity score-based stabilized inverse probability weights (IPTW)
# in the training set to adjust for confounding when modeling treatment effects.
# The target estimand is the ATE
weight_model <- weightit(
  form_denom,
  data = as.data.frame(train),
  stabilize = TRUE,
  method = "ps",
  estimand = "ATE"
)
# Extract stabilized IPTW and store them in the training data
train$sipw_en <- weight_model$weights
```


```{r}
mean(train$sipw_en)

bal.tab(weight_model, un = TRUE)
love.plot(weight_model, binary = "std")
```




```{r}
prep_glinternet <- function(train, test,
                            y_col = "y",
                            cat_cols = character(),
                            tx_col = "tx",
                            drop_cols = NULL) {
  library(dplyr)

  Y_train <- as.integer(train[[y_col]])
  Y_test  <- as.integer(test[[y_col]])

  if (!all(Y_train %in% c(0, 1))) stop("Y_train must be 0/1.")
  if (!all(Y_test  %in% c(0, 1))) stop("Y_test must be 0/1.")

  X_train_df <- train %>% select(-any_of(c(y_col, drop_cols)))
  X_test_df  <- test  %>% select(-any_of(c(y_col, drop_cols)))

  x_cols <- colnames(X_train_df)

  if (tx_col %in% x_cols) {
    X_train_df[[tx_col]] <- as.integer(X_train_df[[tx_col]])
    X_test_df[[tx_col]]  <- as.integer(X_test_df[[tx_col]])
  } else {
    stop("tx column not found in data.")
  }

  cat_levels <- lapply(cat_cols, function(nm) {
    if (!nm %in% x_cols) stop(paste("Categorical var not found:", nm))
    levels(factor(X_train_df[[nm]]))
  })
  names(cat_levels) <- cat_cols

  encode_one <- function(df) {
    X_list <- vector("list", length(x_cols))
    numLevels <- integer(length(x_cols))

    for (j in seq_along(x_cols)) {
      nm <- x_cols[j]
      v  <- df[[nm]]

      if (nm %in% cat_cols) {
        v <- factor(v, levels = cat_levels[[nm]])

        if (any(is.na(v))) {
          stop(paste0(
            "Unseen level in categorical variable '", nm,
            "'. Fix levels BEFORE split."
          ))
        }

        X_list[[j]] <- as.integer(v) - 1L
        numLevels[j] <- length(cat_levels[[nm]])
      } else {
        X_list[[j]] <- as.numeric(v)
        numLevels[j] <- 1L
      }
    }

    X <- do.call(cbind, X_list)
    colnames(X) <- x_cols

    list(X = X, numLevels = numLevels)
  }

  enc_train <- encode_one(X_train_df)
  enc_test  <- encode_one(X_test_df)

  list(
    X_train   = enc_train$X,
    X_test    = enc_test$X,
    Y_train   = Y_train,
    Y_test    = Y_test,
    numLevels = enc_train$numLevels,
    x_cols    = x_cols
  )
}

extract_relaxed_glinternet_terms <- function(cv_fit, x_cols, numLevels, tx_col = "tx") {
  best_idx <- match(cv_fit$lambdaHat, cv_fit$glinternetFit$lambda)
  if (is.na(best_idx)) {
    stop("Could not match lambdaHat to the fitted glinternet path.")
  }

  active <- cv_fit$glinternetFit$activeSet[[best_idx]]

  cont_cols <- x_cols[numLevels == 1]
  cat_cols  <- x_cols[numLevels > 1]

  cont_name <- function(idx) cont_cols[idx]
  cat_name  <- function(idx) cat_cols[idx]

  main_terms <- character()
  int_terms  <- character()

  if (!is.null(active$cont)) {
    main_terms <- c(main_terms, vapply(as.vector(active$cont), cont_name, character(1)))
  }

  if (!is.null(active$cat)) {
    main_terms <- c(main_terms, vapply(as.vector(active$cat), cat_name, character(1)))
  }

  if (!is.null(active$contcont)) {
    contcont_terms <- apply(active$contcont, 1, function(idx) {
      paste(cont_name(idx[1]), cont_name(idx[2]), sep = ":")
    })
    int_terms <- c(int_terms, contcont_terms)
  }

  if (!is.null(active$catcont)) {
    catcont_terms <- apply(active$catcont, 1, function(idx) {
      paste(cat_name(idx[1]), cont_name(idx[2]), sep = ":")
    })
    int_terms <- c(int_terms, catcont_terms)
  }

  if (!is.null(active$catcat)) {
    catcat_terms <- apply(active$catcat, 1, function(idx) {
      paste(cat_name(idx[1]), cat_name(idx[2]), sep = ":")
    })
    int_terms <- c(int_terms, catcat_terms)
  }

  main_terms <- unique(main_terms)
  int_terms  <- unique(int_terms)

  if (!tx_col %in% main_terms) {
    main_terms <- c(main_terms, tx_col)
  }

  list(
    best_idx = best_idx,
    main_terms = unique(main_terms),
    interaction_terms = int_terms
  )
}

build_relaxed_formula <- function(y_col, main_terms, interaction_terms) {
  inherited_main_terms <- unique(unlist(strsplit(interaction_terms, ":", fixed = TRUE)))
  rhs_terms <- unique(c(main_terms, inherited_main_terms, interaction_terms))

  if (length(rhs_terms) == 0) {
    rhs_terms <- "1"
  }

  as.formula(
    paste(y_col, "~", paste(rhs_terms, collapse = " + "))
  )
}

clamp01 <- function(x) {
  pmin(pmax(as.numeric(x), 0), 1)
}
```


# Glinternet estimating 

```{r}
plan(multisession, workers = 6)
pacman::p_load("glinternet", "rsample", "dplyr")
# glinternet: fits hierarchical group-lasso models with interaction selection
# rsample: provides tools for data splitting (train/test)
# dplyr: data manipulation

set.seed(123)
outcome <- "y"
# Reuse the earlier train/test split so weighting, model fitting, and
# downstream prediction all stay aligned to the same partition.

# glinternet requires categorical variables to be explicitly specified
# so that it can correctly handle group structure and interactions
cat_cols <- c("smok")

prep <- prep_glinternet(
  train    = train,
  test     = test,
  y_col    = "y",
  cat_cols = cat_cols,
  tx_col   = "tx",
  drop_cols = c("sipw_en")
)


# Converts data into numeric design matrices (X_train, X_test)
# Creates numLevels (required by glinternet to define group structure)
# Separates predictors (X) and outcome (Y)

# Outputs:
# prep$X_train: predictor matrix for training
# prep$Y_train: outcome vector for training
# prep$X_test : predictor matrix for testing
# prep$numLevels: number of levels for each variable (needed for group lasso)


set.seed(123)
fit_cv <- glinternet.cv(
  X = prep$X_train,
  Y = prep$Y_train,
  numLevels = prep$numLevels,
  family = "binomial",
  nFolds = 10
)




# glinternet.cv:
# Fits a hierarchical group-lasso model
# Automatically selects main effects and pairwise interactions
# Enforces hierarchy: interactions only enter if main effects are included
# Uses cross-validation to choose optimal regularization parameter (lambda)

plot(fit_cv)

relaxed_terms <- extract_relaxed_glinternet_terms(
  cv_fit = fit_cv,
  x_cols = prep$x_cols,
  numLevels = prep$numLevels,
  tx_col = "tx"
)

relaxed_form <- build_relaxed_formula(
  y_col = "y",
  main_terms = relaxed_terms$main_terms,
  interaction_terms = relaxed_terms$interaction_terms
)

relaxed_form

# Stage 2 relaxed refit:
# use a standard weighted logistic regression on the selected terms.
# With continuous IPTW, glm(binomial) may print the usual
# "non-integer #successes" warning from prior weights.
final_model <- glm(
  formula = relaxed_form,
  data = train,
  family = binomial(),
  weights = sipw_en
)

summary(final_model)

p_test <- as.numeric(
  predict(final_model, newdata = test, type = "response")
)

p_test <- clamp01(p_test)

# Observed-treatment prediction only; retained as a quick model check
# and not used for downstream ITE estimation.
head(p_test)


```



```{r}

# Use the fitted cross-validated glinternet model
cv_lasso <- final_model

# Convert test dataset to tibble for easier manipulation
pred_test_df <- test %>% dplyr::as_tibble()
test0 <- pred_test_df %>% dplyr::mutate(tx = 0)
test1 <- pred_test_df %>% dplyr::mutate(tx = 1)

# Predict outcome probability under control (tx = 0)
pred_test_df$y0 <- as.numeric(
  predict(cv_lasso, newdata = test0, type = "response")
)
pred_test_df$y0 <- clamp01(pred_test_df$y0)

# Predict outcome probability under treatment (tx = 1)
pred_test_df$y1 <- as.numeric(
  predict(cv_lasso, newdata = test1, type = "response")
)
pred_test_df$y1 <- clamp01(pred_test_df$y1)

# Compute individual treatment effect (ITE)
pred_test_df$ite <- pred_test_df$y1 - pred_test_df$y0


```


```{r}
# estimate ite in full dataset using the original analysis dataset
df_full_orig <- dsn2 %>% as_tibble()


df_full <- df_full_orig %>% as_tibble()
df_full0 <- df_full %>% dplyr::mutate(tx = 0)
df_full1 <- df_full %>% dplyr::mutate(tx = 1)

df_full$y0 <- as.numeric(
  predict(cv_lasso, newdata = df_full0, type = "response")
)
df_full$y0 <- clamp01(df_full$y0)

df_full$y1 <- as.numeric(
  predict(cv_lasso, newdata = df_full1, type = "response")
)
df_full$y1 <- clamp01(df_full$y1)

df_full$ite <- df_full$y1 - df_full$y0

```




```{r}

# Use glmnet to select important variables for ITE prediction (post-hoc screening)
library(fastDummies)
library(dplyr)

# creat dummy variables for smoking status
smok_df <- fastDummies::dummy_cols(
  data.frame(smok = df_full$smok),
  select_columns = "smok",
  remove_first_dummy = TRUE,        # drop reference
  remove_selected_columns = TRUE
) 

smok_mat <- as.matrix(smok_df)

# Build a numeric design matrix for post-hoc glmnet ITE screening.
# Exclude observed outcome, treatment factor, and all estimated potential
# outcome columns to avoid leakage into the screening model.
X_full_base <- df_full %>%
  dplyr::select(-y, -smok, -y0, -y1, -ite) %>%
  as.matrix()

X_full2 <- cbind(X_full_base, smok_mat)

# ensure tx column is numeric in this design
idx_tx_full2 <- which(colnames(X_full2) == "tx")
if (length(idx_tx_full2) == 1) {
  X_full2[, idx_tx_full2] <- as.numeric(X_full2[, idx_tx_full2])
}

y_ite_scaled <- as.numeric(scale(df_full$ite))
keep <- complete.cases(X_full2) & !is.na(y_ite_scaled)

X_cc <- X_full2[keep, , drop = FALSE]
X_cc <- X_cc[, colnames(X_cc) != "smok", drop = FALSE]
X_cc <- X_cc[, colnames(X_cc) != "tx", drop = FALSE]

y_cc <- y_ite_scaled[keep]

fit_ite_glmnet <- cv.glmnet(
  x = X_cc,
  y = y_cc,
  family = "gaussian",
  alpha  = 1,
  nfolds = 10
)




# Extracting coefficients at the lambda that minimizes cross-validated error

coef_ite <- coef(fit_ite_glmnet, s = "lambda.min")

coef_df <- data.frame(
  variable = rownames(coef_ite),
  coefficient = as.numeric(coef_ite)
) |>
  dplyr::filter(variable != "(Intercept)") |>
  dplyr::arrange(dplyr::desc(abs(coefficient)))
coef_df <- coef_df %>%
  dplyr::mutate(
    coefficient = round(coefficient, 4)
  )
coef_df

```


# plot coeffcient
```{r}
library(dplyr)

coef_df <- coef_df %>%
  mutate(abs_coef = abs(coefficient))

coef_df_nz <- coef_df %>%
  dplyr::filter(abs_coef > 0)


var_labels <- c(
  aspirin     = "Aspirin use",
  egfr        = "Estimated glomerular filtration rate (eGFR)",
  fram        = "Framingham risk score",
  sbp         = "Systolic blood pressure",
  black       = "Black race",
  age         = "Age in years",
  scr         = "Serum creatinine",
  statin      = "Statin use",

  smok_Former = "Former smoker"
  
)


library(ggtext)

ggplot(coef_df_nz, aes(
  x = abs_coef,
  y = reorder(variable, abs_coef)
)) +
  geom_col(fill = "#377EB8") +   
  scale_x_continuous(
    breaks = seq(0, 2, 0.1),
    limits = c(0, max(coef_df_nz$abs_coef))
  ) +
  scale_y_discrete(labels = var_labels) +
  labs(
    x = "Variable importance",
    y = "Predictors",
    title = "Variable importance based on absolute standardized coefficients"
  ) +
  geom_vline(xintercept = 0.2, linetype=2, linewidth=1) +
  theme_bw(base_size = 10) +
  theme(
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.y  = ggtext::element_markdown(face = "bold")
  )


```


```{r}
# fit linear regression using selected variables from glmnet to 
# get standardized coefficients for variable importance plot
# (ite ~ covariates)
sel_vars <- coef_df %>%
  dplyr::filter(coefficient != 0) %>%
  dplyr::pull(variable)


df_X_cc <- as.data.frame(X_cc)


df_lm <- df_X_cc %>%
  dplyr::mutate(ite = y_cc) %>%
  dplyr::select(ite, all_of(sel_vars))


df_lm_scaled <- df_lm %>%
  dplyr::mutate(
    ite = as.numeric(scale(ite))
  ) %>%
  dplyr::mutate(
    dplyr::across(
      all_of(sel_vars),
      ~ as.numeric(scale(.x))
    )
  )

form_lm <- as.formula(
  paste("ite ~", paste(sel_vars, collapse = " + "))
)

fit_lm <- lm(form_lm, data = df_lm_scaled)

summary(fit_lm)



```

# plot lm eatimate
```{r}
library(broom)
library(dplyr)

imp_lm <- broom::tidy(fit_lm) %>%
  filter(term != "(Intercept)") %>%
  mutate(
    estimate = as.numeric(estimate),
    abs_est  = abs(estimate)
  ) %>%
  arrange(desc(abs_est))
library(ggplot2)

library(ggtext)

ggplot(imp_lm, aes(
  x = abs_est,
  y = reorder(term, abs_est)
)) +
  geom_col(fill = "#377EB8") +   
  scale_x_continuous(
    breaks = seq(0, 1, 0.1),
    limits = c(0, max(imp_lm$abs_est))
  ) +
  scale_y_discrete(labels = var_labels) +
  labs(
    x = "Variable importance",
    y = "Predictors",
    title = "Variable importance based on absolute standardized coefficients"
  ) +
  geom_vline(xintercept = 0.2, linetype=2, linewidth=1) +
  theme_bw(base_size = 13) +
  theme(
    axis.title.x = element_text(face = "bold"),
    axis.title.y = element_text(face = "bold"),
    axis.text.y  = ggtext::element_markdown(face = "bold")
  )


```



# Histogram and Ranking of Estimated ITE

```{r}
# Function to plot histogram as percentages
geom_hist_percent <- function(.data, .var) {
  .data %>%
    ggplot(aes(x = !!sym(.var))) +
    geom_histogram(aes(y = after_stat(count/sum(count)))) +
    ylab("Percent") +
    scale_y_continuous(labels = scales::percent) +
    theme_bw()
}

# Histogram of ITE distribution
pred_test_df %>%
  geom_hist_percent("ite") +
  labs(x="Individualized treatement effects (ITE)") 

# Rank plot of ITE values
pred_test_df %>%
  arrange(ite) %>%
  mutate(rank=1:n()) %>%
  ggplot(aes(y=ite, x=rank)) +
  geom_point() +
  labs(y="Individualized effects (difference scale)") +
  geom_hline(yintercept = 0, linetype=2) +
  theme_bw()

```


# Density Plot of ITE (Percentage Scale)
```{r}
# Convert ITE to percentage points and plot density
pred_test_df %>%
  mutate(ite_pp = ite * 100) %>%
  ggplot(aes(x = ite_pp)) +
  geom_density(fill = "steelblue", alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = 2) +
  labs(x = "Estimated ITE (percentage points)") +
  theme_bw()

```

```{r}
pred_test_df <- pred_test_df %>%
  mutate(
    smok_2 = as.integer(smok == "Former"),
    smok_3 = as.integer(smok == "Current")
  )%>%
  select(-smok)
```

# Compute Probabilities and Scaled ITE

```{r}


pred_test_df$ite <- pred_test_df$y1 - pred_test_df$y0
pred_test_df$ite_scaled <- as.numeric(scale(pred_test_df$ite
                                          )) #--Scale for clustering


# Add Noise and Prepare for Clustering
set.seed(123)
pred_test_df <-  pred_test_df %>% 
  mutate(
    iter = ite + rnorm(n=n(), mean=0, sd=0.1), #jittering
    iter_scaled = as.numeric(scale(iter)))

# Determine Optimal Number of Clusters
set.seed(123)
# Use NbClust to determine optimal number of clusters
p_load("NbClust")
res<-NbClust(as.data.frame(pred_test_df$iter_scaled), 
             distance = "euclidean", 
             min.nc=2, max.nc=15, 
             method = "ward.D2", 
             index = "beale", alphaBeale = 0.05)

num_cluster <- as.numeric(res$Best.nc[1])
num_cluster

```

```{r}
# Assign Clusters and Update Covariates
pred_test_df$cluster <- res$Best.partition
covariates <- setdiff(colnames(pred_test_df), c("y", "tx"))
covariates <- setdiff(covariates, exposure)
covariates <- setdiff(
  covariates,
  c("y0", "y1", "ite", "ite_scaled", "iter", "iter_scaled", "cluster")
)
```

```{r}
library(ggtext)
# Compute cluster-level means for ITE and covariates
tr <- pred_test_df %>%
  dplyr::select(-all_of(exposure)) %>% 
  group_by(cluster) %>%
  group_modify(~ {.x %>% map_dfr(mean)}) %>%
  ungroup() %>% 
  arrange(cluster, ite_scaled) %>% 
  relocate(cluster, ite_scaled) %>% 
  dplyr::select(cluster, ite, ite_scaled, all_of(covariates)) 

# Select variables with meaningful differences across clusters
varnames <- pred_test_df %>%
  dplyr::select(-all_of(exposure)) %>% 
  mutate_at(covariates, \(x) as.numeric(scale(x))) %>% 
  group_by(cluster) %>%
  group_modify(~ {.x %>% map_dfr(mean)}) %>%
  ungroup() %>%
  dplyr::select(cluster, all_of(covariates)) %>% 
  mutate_all(diff_fct) %>% 
  dplyr::select(-where(~any(is.nan(.)))) %>% 
  drop_na() %>% 
  dplyr::select(where(~mean(abs(.x), na.rm=T) > 0.2)) %>% 
  dplyr::select(-cluster) %>% 
  colnames()

# Calculate variable importance based on between-cluster differences
var_imp <- pred_test_df %>%
  dplyr::select(-all_of(exposure)) %>% 
  mutate_at(covariates, \(x) as.numeric(scale(x))) %>% 
  group_by(cluster) %>%
  group_modify(~ {.x %>% map_dfr(mean)}) %>%
  ungroup() %>%
  dplyr::select(all_of(covariates)) %>% 
  mutate_all(diff_fct) %>% 
  dplyr::select(-where(~any(is.nan(.)))) %>% 
  mutate_all(abs) %>% 
  drop_na() %>% 
  pivot_longer(cols=everything(),
               names_to = "variables",
               values_to = "importance") %>% 
  arrange(desc(importance))


cb <- data.frame(labels=var_label(dsn1) %>% unlist()) %>% 
  rownames_to_column(var="variables") %>% 
  filter(variables %in% var_imp$variables)

# Merge labels with importance results
var_imp_labels <- var_imp %>% 
  left_join(cb, by="variables")
# Plot variable importance
var_imp_plot <- var_imp_labels %>%
  mutate(labels = str_wrap(labels, width = 10)) %>% 
  ggplot(aes(y=reorder(labels, importance), x=importance)) +
  geom_bar(stat= "identity") +
  geom_vline(xintercept = 0.2, linetype=2, linewidth=1) +
  scale_fill_brewer(palette="Set1") +
  scale_x_continuous(breaks=seq(0,1.5,0.1)) +
  ylab("Predictors") + 
  xlab("Variable importance") +
  theme_bw() +
  theme(axis.title.x = element_text(face = "bold"),
        axis.text.y =  ggtext::element_markdown(face = "bold"))

# Create summary table of important variables by cluster
var_imp_cut <- pred_test_df %>%
  dplyr::select(cluster, ite, all_of(varnames)) %>% 
  group_by(cluster) %>%
  group_modify(~ {.x %>% map_dfr(mean)}) %>%
  ungroup() %>%
  arrange(cluster, ite) %>% 
  round_df(2)


```

```{r}

var_imp_cut %>% flextable()
```

```{r}
varnames

var_imp_plot
```


```{r}
# Load emmeans for marginal effects
p_load("emmeans")

dsn_final <- dsn1
covariates
```
# Propensity Score Weighting (IPTW)
```{r}
form_denom <- tx ~ age + black + sbp + dbp + scr + egfr + statin + aspirin + 
    fram + smok_2 + smok_3
weight_model_all <- weightit(form_denom, 
                         data = dsn_final, 
                         stabilize = T,
                         method = "ps", 
                         estimand = "ATE")

# Attach weights and create ID
dsn_final$sipw_en_all <- weight_model_all$weights
dsn_final1 <- dsn_final %>% 
  mutate(id=1:n())

#check the weights
mean(dsn_final$sipw_en_all)

hist(dsn_final$sipw_en_all)

# Covariate balance plot
love.plot(weight_model_all, 
          var.order = "unadjusted",
          abs = TRUE,              
          threshold = 0.1)   


```

# Estimate ATE Using Weighted Regression
```{r}
pacman::p_load("lmtest", "sandwich")

# Fit weighted model (risk difference scale)
ate_rd <- glm(y~tx, 
              family=quasibinomial("identity"), 
              data=dsn_final1,
              weights = sipw_en_all)
# Compute robust clustered standard errors
dn_ate_rd <-coeftest(ate_rd, vcov = vcovCL(ate_rd, cluster=dsn_final1$id, type="HC1")) %>% 
  tidy(conf.int=T) %>% 
  filter(term=="tx") %>% 
  dplyr::select(term,estimate,conf.low, conf.high,p.value) %>% 
  round_df(2) 

dn_ate_rd %>% 
  flextable()
```

```{r}
# Fit weighted model (risk ratio scale)
ate_rr <- glm(y~tx, 
              family=quasibinomial("log"), 
              data=dsn_final1,
              weights = sipw_en_all)

dn_ate_rr <- coeftest(ate_rr, vcov = vcovCL(ate_rr, cluster=dsn_final1$id, type="HC1")) %>% 
  tidy(conf.int=T) %>% 
  filter(term=="tx") %>% 
  mutate_at(c("estimate","conf.low", "conf.high"), exp) %>%
  dplyr::select(term,estimate,conf.low, conf.high,p.value) %>% 
  round_df(2)

    
dn_ate_rr %>% 
  flextable()
```

# Conditional average treatment effect (CATE)
```{r}
varnames

var_imp_cut
```

```{r}
cut_egfr <- round(mean(c(var_imp_cut$egfr[1], var_imp_cut$egfr[2])),2)
cut_egfr
```

## eGFR Risk difference
```{r}
dsn_final2 <- dsn_final1 %>%
  mutate(egfr_bin = as.numeric(round(egfr>=cut_egfr)))


### eGFR Risk difference
model_cate_egfr_rd <- glm(y~tx*egfr_bin,
              family=quasibinomial("identity"),
              data=dsn_final2,
              weights = sipw_en_all)

tidy(model_cate_egfr_rd) %>% 
  filter(term=="tx:egfr_bin") %>% 
  select(term, `p.value`) %>% 
  round_df(4)
# check out the p-value for the interaction term# check out the p-value for the interaction term


p_load("emmeans")
cate_egfr_rd <- emmeans::emtrends(model_cate_egfr_rd, ~egfr_bin,
                          var="tx",
                          at = list(egfr_bin  =c(0,1)),
                          vcov = sandwich::vcovCL(model_cate_egfr_rd,
                                                  cluster= dsn_final2$id,
                                                  type="HC1")) %>%
  tidy(conf.int=T) %>%
  round_df(2) %>%
  flextable()
cate_egfr_rd

```

```{r}
### eGFR Risk ratio
model_cate_egfr_rr <- glm(y~tx*egfr_bin,
                          family=quasibinomial("log"),
                          data=dsn_final2,
                          weights = sipw_en_all)

tidy(model_cate_egfr_rr)  %>% 
  filter(term=="tx:egfr_bin") %>% 
  select(term, `p.value`) %>% 
  round_df(4)# check out the p-value for the interaction term

p_load("emmeans")
cate_egfr_rr <- emmeans::emtrends(model_cate_egfr_rr, ~egfr_bin,
                                  var="tx",
                                  at = list(egfr_bin  =c(0,1)),
                                  vcov = sandwich::vcovCL(model_cate_egfr_rr,
                                                          cluster= dsn_final2$id,
                                                          type="HC1")) %>%
  tidy(conf.int=T) %>%
  mutate_at(c("tx.trend", "conf.low",  "conf.high"), exp) %>%
  round_df(2) %>%
  flextable()
cate_egfr_rr
```

```{r}
# Number count and proportion
library(dplyr)
library(tibble)


N_prop_eGFR <- dsn_final2  %>%
  filter(!is.na(y), !is.na(tx), !is.na(egfr_bin)) %>%
  count(egfr_bin, name = "N") %>%
  mutate(
    Proportion = round(100 * N / sum(N), 2)
  ) %>%
  arrange(egfr_bin)

N_prop_eGFR


```




## Aspirin
```{r}

## aspirin
cut_aspirin <- round(mean(c(var_imp_cut$aspirin[1], var_imp_cut$aspirin[2])),2)
cut_aspirin

dsn_final2 <- dsn_final1 %>%
  mutate(aspirin_bin = as.numeric(round(aspirin>=cut_aspirin)))


### aspirin Risk difference
model_cate_aspirin_rd <- glm(y~tx*aspirin_bin,
                             family=quasibinomial("identity"),
                             data=dsn_final2,
                             weights = sipw_en_all)

tidy(model_cate_aspirin_rd)  %>% 
  filter(term=="tx:aspirin_bin") %>% 
  select(term, `p.value`) %>% 
  round_df(4)# check out the p-value for the interaction term
p_load("emmeans")
cate_aspirin_rd <- emmeans::emtrends(model_cate_aspirin_rd, ~aspirin_bin,
                                     var="tx",
                                     at = list(aspirin_bin  =c(0,1)),
                                     vcov = sandwich::vcovCL(model_cate_aspirin_rd,
                                                             cluster= dsn_final2$id,
                                                             type="HC1")) %>%
  tidy(conf.int=T) %>%
  round_df(2) %>%
  flextable()
cate_aspirin_rd
```

```{r}
### aspirin Risk ratio
model_cate_aspirin_rr <- glm(y~tx*aspirin_bin,
                             family=quasibinomial("log"),
                             data=dsn_final2,
                             weights = sipw_en_all)

tidy(model_cate_aspirin_rr) %>% 
  filter(term=="tx:aspirin_bin") %>% 
  select(term, `p.value`) %>% 
  round_df(4)

# check out the p-value for the interaction term

p_load("emmeans")
cate_aspirin_rr <- emmeans::emtrends(model_cate_aspirin_rr, ~aspirin_bin,
                                     var="tx",
                                     at = list(aspirin_bin  =c(0,1)),
                                     vcov = sandwich::vcovCL(model_cate_aspirin_rr,
                                                             cluster= dsn_final2$id,
                                                             type="HC1")) %>%
  tidy(conf.int=T) %>%
  mutate_at(c("tx.trend", "conf.low",  "conf.high"), exp) %>%
  round_df(2) %>%
  flextable()

cate_aspirin_rr
```


```{r}
# Number count and proportion
library(dplyr)
library(tibble)


N_prop_aspirin <- dsn_final2  %>%
  filter(!is.na(y), !is.na(tx), !is.na(aspirin_bin)) %>%
  count(aspirin_bin, name = "N") %>%
  mutate(
    Proportion = round(100 * N / sum(N), 2)
  ) %>%
  arrange(aspirin_bin)

N_prop_aspirin


```







# plot Aspirin and eGFR HTE

```{r}
library(ggplot2)
library(dplyr)

hte_cut <- 0

df_lasso_plot <- df_full %>%
  mutate(
    ite_pct = ite * 100,
    aspirin_f = factor(aspirin, levels = c(0, 1),
                       labels = c("No Aspirin", "Aspirin"))
  ) %>%
  filter(!is.na(aspirin_f), !is.na(ite_pct))

ggplot(df_lasso_plot, aes(x = aspirin_f, y = ite_pct)) +
  geom_jitter(
    width = 0.15,
    height = 0,
    alpha = 0.55,
    size = 1.8,
    color = "#4C72B0"
  ) +
  geom_hline(
    yintercept = hte_cut,
    linetype = "dashed",
    linewidth = 1,
    color = "#D7191C"
  ) +
  labs(
    x = "Baseline Aspirin Use",
    y = "Estimated HTE (percentage points)"
  ) +
  theme_bw(base_size = 14) +
  theme(
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "grey92", linewidth = 0.8),
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold"),
    axis.text  = element_text(color = "black")
  )
library(dplyr)

quad_counts <- df_lasso_plot %>%
  mutate(
    quad = case_when(
      aspirin == 0 & ite_pct >= 0 ~ "Q1_left_up",
      aspirin == 0 & ite_pct <  0 ~ "Q3_left_low",
      aspirin == 1 & ite_pct >= 0 ~ "Q2_right_up",
      aspirin == 1 & ite_pct <  0 ~ "Q4_right_low"
    )
  ) %>%
  count(quad)
quad_counts

```


```{r}
library(ggplot2)
library(dplyr)

aspirin_bin <- 0.5
hte_cut   <- 0

df_lasso_plot <- df_full %>%
  mutate(
    ite_pct = ite * 100,
    aspirin_grp = factor(
      ifelse(aspirin >= 0.5, "Aspirin", "No Aspirin"),
      levels = c("No Aspirin", "Aspirin")
    ),
    quad = case_when(
      egfr <  cut_egfr & ite_pct >= hte_cut ~ "Q1_left_up",
      egfr >= cut_egfr & ite_pct >= hte_cut ~ "Q2_right_up",
      egfr <  cut_egfr & ite_pct <  hte_cut ~ "Q3_left_low",
      egfr >= cut_egfr & ite_pct <  hte_cut ~ "Q4_right_low"
    )
  ) %>%
  filter(!is.na(egfr), !is.na(ite_pct), !is.na(aspirin_grp))
quad_cols <- c(
  "Q1_left_up"  = "#8E87E8",
  "Q2_right_up" = "#F28E8C",
  "Q3_left_low" = "#8FD3B6",
  "Q4_right_low"= "#9AD0EC"
)


ggplot(df_lasso_plot,
       aes(x = egfr, y = ite_pct,
           color = quad,
           shape = factor(aspirin))) +
  geom_point(alpha = 0.5, size = 1.6) +
  geom_vline(xintercept = cut_egfr, linetype = "dashed") +
  geom_hline(yintercept = hte_cut, linetype = "dashed") +
  scale_color_manual(values = quad_cols, guide = "none") +
  scale_shape_manual(
    values = c("0" = 16, "1" = 17),
    labels = c("No Aspirin", "Aspirin")
  ) +
  labs(
    x = "Baseline eGFR",
    y = "Estimated HTE (percentage points)",
    shape = "Aspirin use"
  ) +
  theme_bw(base_size = 14)

```


```{r}
# Calculating the number of observations in each quadrant
quad_N <- df_lasso_plot %>%
  count(quad)
quad_N
```


```{r}
# spearman correlation between eGFR and ITE
cor_test_ite <- cor.test(
 df_full$egfr,
  df_full$ite,
  method = "spearman"
)

cor_test_ite


```


```{r}
# Spearman correlation between aspirin and ITE
cor_test_ite <- cor.test(
 df_full$aspirin,
  df_full$ite,
  method = "spearman"
)
cor_test_ite
```
